Characteristic polynomials of p-adic matrices

### 2017-01-25 charpoly/charpoly.tex
Added abstract, working on Section 2

\title{Characteristic polynomial of p-adic matrices}

@@

\begin{abstract}
\end{abstract}

@@

The main benefit of working with lattices is the following result
\citelist{\cite{caruso-roe-vaccon:14a}*{Lem. 3.4} \cite{caruso-roe-vaccon:15a}*{Prop. 2.2}}.

%\begin{prop}

%\end{prop}

\todo{Recall the Lemma (with integral version)}

@@

@@

\section{Computation of $\chi$ and $d \chi$}

In this section, we combine the computation of an Hessenberg form
of a matrix and the computation of the inverse through the SNF
over a complete discrete valuation field
to compute $\com (X-A)$ and $d \chi.$
%The computation of an Hessenberg form provides an $O(n^3)$
%algorithm to compute the characteristic polynomial.

>>>>

\title{Characteristic polynomials of p-adic matrices}

@@

\begin{abstract}
We analyze the precision of the characteristic polynomial $\chi(A)$ of
an $n \times n$ $p$-adic matrix $A$ using differential precision methods
developed previously.  When $A$ is integral with precision $O(p^k)$,
we give a criterion (checkable in time $\softO(n^\omega)$) for
$\chi(A)$ to have precision exactly $O(p^k)$.  We also give a $\softO(n^3)$
algorithm for determining the optimal precision when the criterion is not
satisfied, and give examples when the precision is larger than $O(p^k)$.
\end{abstract}

@@

Let $\chi : M_n(K) \to K_n[X]$ be the characteristic polynomial map.  For a matrix $M \in M_n(K)$
we identify the tangent space $V$ at $M$ with $M_n(K)$ itself, and the tangent space $W$ at $\chi(M)$ with
the space $K_{<n}[X]$ of polynomials of degree less than $n$.  Write $B^-_V(r)$ (resp. $B_V(r)$) for the open (resp. closed)
ball of radius $r$ in $V$.  
Our analysis of the $\chi$ begins with the following result
\citelist{\cite{caruso-roe-vaccon:14a}*{Lem. 3.4} \cite{caruso-roe-vaccon:15a}*{Prop. 2.2}}.

\begin{prop} \label{prop:mainlem}
Suppose that $M \in M_n(K)$ and that $d\chi(M): V \to W$ is surjective.
Let $C$ be a positive real number such that
$B_W(1) \subset d\chi(M)(B_V(C))$.  Then, for all $\rho \in (0, 1]$ and all
$r \in (0, C \cdot \rho^{-1})$, any lattice $H$ such that $B_V^-(\rho r) \subset H \subset B_V(r)$
satisfies:
\begin{equation}
\chi(M + H) = \chi(M) + d\chi(H).
\end{equation}
\end{prop}

@@

In light of Proposition \ref{prop:mainlem}, we recall the computation of $d\chi$
\citelist{\cite{caruso-roe-vaccon:14a}*{Appendix B} \cite{caruso-roe-vaccon:15a}*{\S 3.3}}.
Let $\com(M)$ denote the comatrix of $M$; when $M \in \GL_n(K)$ then $\com(M) = \det(M) M^{-1}$.  Then $d\chi(M)$ is given
by
\[
d\chi(M): dM \mapsto \tr(\com(X-M) \cdot dM).
\]
Moreover, $d\chi(M)$ will be surjective exactly when the Jordan blocks of $M$ all have distinct generalized eigenvalues.

@@

Equality of lattices holds iff it does mod p, so need surjectivity modulo p.  Surjectivity of differential
can be seen: write matrix in Jordan normal form.  Comatrix(X - M) surjective iff when you write it
in Frobenius normal form it has only one block.

\section{Computation of the characteristic polynomial and its derivative}

In this section, we combine the computation of an Hessenberg form
of a matrix and the computation of the inverse through the SNF
over a complete discrete valuation field
to compute $\com (X-M)$ and $d \chi.$
%The computation of an Hessenberg form provides an $O(n^3)$
%algorithm to compute the characteristic polynomial.
